**2.5**

1.  **Why does this code not work?**

    a változó my_variable, a következő sorban my_varıable van

    ```{r}
    my_variable <- 10
    my_varıable
    ```

2.  **Tweak each of the following R commands so that they run correctly:**

    ```         
    libary(todyverse)
    ```

    ```{r}
    library(tidyverse)
    ```

    ```         
    ggplot(dTA = mpg) +
      geom_point(maping = aes(x = displ y = hwy)) +
      geom_smooth(method = "lm)
    ```

    ```{r}
    ggplot(
      data = mpg,
      mapping = aes(x = displ, y = hwy)) +
      geom_point() +
      geom_smooth(method = "lm")
    ```

3.  **Press Option + Shift + K / Alt + Shift + K. What happens? How can you get to the same place using the menus?**

    megjelenik a shortcutok listája\
    help -\> keyboard shortcuts help

4.  **Run the following lines of code. Which of the two plots is saved as `mpg-plot.png`? Why?**

    a bar plot, mivel a plot argumentum ezt határozta meg

    ```{r}
    my_bar_plot <- ggplot(mpg, aes(x = class)) +
      geom_bar()
    my_scatter_plot <- ggplot(mpg, aes(x = cty, y = hwy)) +
      geom_point()
    ggsave(filename = "mpg-plot.png", plot = my_bar_plot)
    ```
